

Age estimation is one of the most relevant tasks in the field of forensic anthropology and is useful both for the identification of living and dead persons in problems as different as disappearances of persons, migratory crises, natural catastrophes, unsolved crimes, etc. 
Among the various existing techniques for estimating the age of death in deceased individuals, the study of the morphological traits present in the pubic symphysis, a method proposed by T. W. Todd in 1921 and still used today, stands out. This method presents a high level of subjectivity, since the study of these characteristics depends almost entirely on the experience of the forensic anthropologist. Therefore, the low level of application of existing automatic techniques to support this task stands out.


This TFG addresses the development of a system capable of estimating the age of an individual from a 3D model of the pubic symphysis. Our proposal presents a totally new and transgressive approach with respect to current techniques. Unlike these, where the participation of an expert forensic anthropologist is necessary to extract the morphological features of the samples, the method we propose will be able to extract these features automatically. We achieve this goal by approaching this problem within the deep learning paradigm. In order to apply techniques related to this field we use a panoramic projection method that allows us to transform 3D models into 2D images while maintaining the morphological information of the models. To train the proposed deep convolutional models we used the set of 1104 3D pubic symphysis models owned by the University of Granada. This set has the limitations of being too limited in number and having highly unbalanced classes, so we have applied data augmentation and balancing strategies.

The results obtained by the proposed method are comparable to those of the current state of the art, so the obtained model is highly competitive and presents a viable alternative to support the age estimation of an individual.